PhD Title \\ PhD Thesis

### 2024-07-11 Latex_files/Thesis/chapters/chapter 2/main.tex
updates

The force exerted by an optical tweezer can be subdivided into the gradient and scattering components, for most modelling research this is how the force fields are reported. The gradient force is a conservative force brought about by the polarisation of dielectric materials, anything around 10\% of the trapping wavelength will be attracted along the gradient of the electric field to the region of greatest intensity (for a simple Gaussian Beam this would be at the focal point). The scattering force arises from the scattered field of the trapping beam pushing the target object away from the focal point; unlike the gradient force, the scattering force is non-conservative and is loosely proportional to the particle's Brownian motion. The equilibrium position (where the mean squared displacement is minimised) is found when the gradient force far exceeds the scattering force, interpreting and calculating these forces is dependent on the ratio between the particle size and the trapping wavelength.

@@

The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align} 
Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$) This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:
\begin{align}
  {\bf E}_{\rm inc}({\bf r})
  &=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}
where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, for the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user.  However, computing the expansion coefficients of the
internal and external fields be far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- all of which is discussed later on.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields, essentially
computing the momentum contained in the incident beam and how much of
it has been transferred to the target particle.

@@

The Rayleigh approximation is the limiting approximation for describing a particles motion in an electromagnetic field who's wavelength is several times greater than the particle's size. The underlying theory is that a dielectric sphere can be treated as a dipole while in the presence of the electromagnetic field; in which case the scattering force is given simply by the scattering of the induced dipole, and the gradient force is due to the Lorentz force \cite{Gordon1973}. The gradient forces in the principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

The Rayleigh regime allows for easy computation of the gradient and scattering forces due to the assumption that the particle is a point dipole, so much so that higher order scattering problems can be simplified by subdividing the particle into discrete dipoles (see Sec \ref{sec:scattering}). However as the target particle gets larger this assumption fails to accurately describe the trapping force due to the complexity in the gradient field. For particle sizes close to the laser wavelength the scattered field is best described by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Tmatrix}

The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering, this would later be extended
to electromagnetic waves. Put simply, the method sets a boundary
condition where the incident field within the interior field is
cancelled out via interference; these boundary conditions allows us to
solve the Maxwell equations by expanding the incident and scattered
fields using spherical vector waves. Because the boundary condition
ensures that the incident field is halted at the interior surface of
the object, the scattering field can be related to the incident field
by the objects $T$-matrix:
\begin{align}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

The $T$-matrix method is exceptionally useful for computing the
scattering from any arbitrary shaped particle.  However, the
$T$-matrix method by itself can be limited by computation power for
aggregates of particles; while possible to compute the scattering of
each individual particle the system of equations cannot be solved
analytically and must be solved numerically to determine properties
for the aggregate for one particular orientation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\subsection{Discrete Dipole Approximation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules diffusive behaviour, for the case of optical tweezers the most complete model of diffusion is the Langevin equation. Models such as the Fickian, and Einstein derivations are sufficient for macroscopic behaviours the Langevin equation better describes the microscopic characteristics of any diffusive behaviour. The Fickian model is not suitable for the applications of optical tweezers as it assumes that all diffusive motion is driven by a gradient of molecular density, not only is this often not the case but it fails to describe the motion of any single molecule, only the collective behaviour. Einstein's model however considers the collisions experienced by individual molecules, if we assume that all collisions are random and only consider the behaviour after a finite time step $\delta t$ then the individual collisions should cancel out. This is insufficient for our interests as an optical tweezers relaxation time (given by $\tau = \kappa/\gamma$) is often so short that we must consider the individual collisions experienced by any one molecule. The Langevin model of diffusion therefore assumes that the net force on a particular particle is described fully by these individual collisions:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}
Where the first term accounts for inertial forces, the second term accounts for friction forces which counteract the particles current motion ($\gamma_0$ is the friction coefficient), and the final term accounts for the random Brownian force. The $F(t)$ is there for convention which accounts for any external forces acting on the particle. We can say that the noise term $W(t)$ has a Gaussian probability, with a correlation function of:

\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

@@

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial elements in order to approximate the higher order terms. If we assume that $x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}
Where $R_n(x(t))$ is the remainder term between the Taylor expansion to term n and the actual expression. If we limit our approach to the first derivative only we find that for sufficiently small values of $R_1$ the velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}
By reversing the time step (i.e. use $-\Delta t$) to approximate the velocity and acceleration based on the previous time steps we can discretize the position by taking finitely small  time steps (i.e. $x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for white noise as no information is known about $W(t)$ at any time. We can instead say that the velocity of a Brownian particle should approximate our white noise as a random walk, where at each new time step the position changes randomly within a given range. Constricting the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}
Where $w_i$ is a random real number between -1 and 1, we can say that it is normally distributed around 0 for simulation purposes. We can rearrange this for $x_i$ to approximate the Brownian motion of a particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. Now in the case of an optical trap the restoration time scale is given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles is far greater than the characteristic time, therefore for simulation purposes we can neglect the particle's inertia which allows us to write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

@@

\subsection{Potential Well Analysis}
The Langevin equation for an optically trapped assumes that the trap acts similar to a Hookean spring that creates a potential well about its center. Therefore a simple analysis method is to understand the height and width of said potential well.

Potential analysis is a useful technique for estimating the strength of an optical trap; this method assumes that the force acting on the particle is purely conservative, an accurate presupposition if we ignore the motion of the particle as it enters the trap. This is because the scattering force is far more significant far away from the potential well and is negligible if the trap strength is much greater than the thermal fluctuations. With this in mind we can write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{B}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able to describe the probability distribution $\rho(x)$ centred at some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}
By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} we can fit the potential well in order to determine the trap strength $\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. This has some limitations in that the large fluctuations can throw off the fit meaning a longer acquisition time is necessary to properly fit the potential well, making it difficult to characterise weakly trapped particles who may not remain trapped for long. It also provides no information on the particle itself (i.e. the friction coefficient $\gamma$ and diffusion coefficient $D$).

@@

By taking a time average over multiple trajectories to get $\left<x-x_{eq}\right>$ we can get a fairly accurate estimation of the trap strength. Because of this requires a time average of the particle's displacement any large errors in the position measurement can have knock-on effects. Likewise with the potential-analysis route, no information is gleaned about the particle itself.

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing the random motion of a given particle (or group of particles). This is useful information if say for example we want to understand reaction kinetics on the surface of a catalyst, if we know how far its likely to move from the surface we can tell if its likely to react when a catalytic site becomes available. As it pertains to colloids, consider a suspension of silica spheres immersed in a fluid undergoing Brownian motion (as described by the Langevin Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}
Where $\gamma$ is the objects friction coefficient which for spheres is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white noise variable that is directly related to the thermal energy of the surrounding fluid. If the motion is truly random then we should see an average displacement of 0 regardless of how long we measure for. If we wish to understand the effects of a given external factor (i.e. an electric field or localised heating), simply looking at displacement will reveal nothing of value as its difficult to differentiate between diffusive and a biased motion.

>>>>

The force exerted by an optical tweezer can be subdivided into the 
gradient and scattering components, for most modelling research this 
is how the force fields are reported. The gradient force is a conservative 
force brought about by the polarisation of dielectric materials, 
anything around 10\% of the trapping wavelength will be attracted along 
the gradient of the electric field to the region of greatest intensity 
(for a simple Gaussian Beam this would be at the focal point). The 
scattering force arises from the scattered field of the trapping beam 
pushing the target object away from the focal point; unlike the gradient \\
force, the scattering force is non-conservative and is loosely 
proportional to the particle's Brownian motion. The equilibrium position 
(where the mean squared displacement is minimised) is found when the 
gradient force far exceeds the scattering force, interpreting and 
calculating these forces is dependent on the ratio between the particle 
size and the trapping wavelength.

@@

The Lorenz Mie theory provides an exact solution to the maxwell equations 
for the scattering caused by an isotropic sphere. The theory describes the 
scattered wave given off by a dielectric sphere when incident by a plane 
wave as a summation of partial spherical waves. For any spherical wave the 
vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align}

Where $k$ is the wave number of the electromagnetic radiation ($k = 2\pi/\lambda$)
 This combined with the constraints of Maxwell's Equations leaves 
 very few exact solutions apart from spherical or planar waves, 
 both of which can be converted between with relative ease. For 
 example, a plane wave electric field can be expanded into spherical 
 harmonics and likewise any spherical wave can be described as 
 combination of plane waves offset from one another. For a single 
 plane wave the incident, internal, and scattered fields are given as:
\begin{align}
	\label{eq:lornez_mie}
  {\bf E}_{\rm inc}({\bf r})
  &=
    E_0 \sum_{nm}^\infty \left[
    a_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    + b_{nm}{\bf N}_{nm}^{(1)}({\bf r})\right]
  \\
  \label{eq:internal}
  {\bf E}_{\rm int}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty i^n\frac{2n+1}{n(n+1)}\left[
    -id_{nm}{\bf N}_{nm}^{(1)}({\bf r}) + c_{nm}{\bf M}_{nm}^{(1)}({\bf r})
    \right]
  \\
  {\bf E}_{\rm scat}({\bf r})
  &=
    E_0 \sum_{n=1}^\infty  i^n\frac{2n+1}{n(n+1)}\left[
      -i p_{nm}{\bf N}_{nm}^{(3)}({\bf r})+q_{nm}{\bf M}_{nm}^{(3)}({\bf r})
    \right] 
    \\
	\mathbf{E}_{tot}({\bf r}) &= 
	\begin{cases}
		{\bf E}_{\rm inc}({\bf r}) + {\bf E}_{\rm scat}({\bf r}) & \text{if \textbf{r} outside}
		\\ 
		{\bf E}_{\rm int}({\bf r}) & \text{if \textbf{r} inside}
	\end{cases}
\end{align}
where $a_{nm}$, $b_{nm}$, $c_{nm}$, $d_{nm}$, $p_{nm}$, and $q_{nm}$
are the expansion coefficients of each of the fields, and $M^{(1)}_{nm}$ 
and $N^{(1)}_{nm}$ are the vector spherical harmonics. For the incident
field computing its expansion coefficients is possible via analytical
methods and are completely dependent on the beam conditions imposed by
the user. However, computing the expansion coefficients of the
internal and external fields be far more tedious depending on the
shape of the target and what properties of the scattered field are
desired --- all of which is discussed later on.  From an optical
trapping perspective the force exerted by a focused electric field can
be found by computing the Maxwell stress tensor which only requires
the total magnitude of the incident and scattered fields, essentially
computing the momentum contained in the incident beam and how much of
it has been transferred to the target particle.

@@

The Rayleigh approximation is the limiting approximation for describing 
a particles motion in an electromagnetic field who's wavelength is several 
times greater than the particle's size. The underlying theory is that a \\
dielectric sphere can be treated as a dipole while in the presence of the 
electromagnetic field; in which case the scattering force is given simply 
by the scattering of the induced dipole, and the gradient force is due 
to the Lorentz force \cite{Gordon1973}. The gradient forces in the 
principle Cartesian axis are described by Harada et al \cite{YasuhiroHarada1996} 
in MKS units as a restorative rectangular force field:
\begin{align}
  F_{grad,x}
  &=-\hat{x} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{x}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,y}
  &=-\hat{y} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2} \times I(r)
  \\
  F_{grad,z}
  &=-\hat{z} \frac{2\pi n_2 a^3}{c}
    \left(\frac{m^2-1}{m^2+2}\right) \frac{4\tilde{y}/w_0}{1+(2\tilde{z})^2}
    \nonumber \times I(r)
  \\ 
  & \times \left[1-\frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
  \\
  \text{where:}
  \nonumber
  \\
	I(r) &= \left(\frac{2P}{\pi w_0^2}\right) \frac{1}{1+(2\tilde{z}^2)} 
	\exp \left[ - \frac{2(\tilde{x}^2+\tilde{y}^2)}{1+(2\tilde{z})^2} \right]
\end{align}

@@

The Rayleigh regime allows for easy computation of the gradient and 
scattering forces due to the assumption that the particle is a point 
dipole, so much so that higher order scattering problems can be 
simplified by subdividing the particle into discrete dipoles (see 
Sec~\ref{sec:scattering}). However as the target particle gets larger 
this assumption fails to accurately describe the trapping force due 
to the complexity in the gradient field. For particle sizes close to 
the laser wavelength the scattered field is best described by Lorenz-Mie theory.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Scattering methods}
\label{sec:scattering}
There are several methods available to calculate the scattered field 
produced by a particle, while one can compute this from directly applying
Lorenz-Mie theory for complicated particle shapes and focused beams makes it 
tedious for some applications.  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{T-matrix Method}
The $T$-matrix method was first developed by Peter Waterman with his
research into acoustic wave scattering \cite{Waterman1969}, this 
would later be extended to electromagnetic waves. Sometimes being 
referred to as the extended boundary condition method (ECBM), 
the method replaces the scatterer with a series of surface currents
over the targets surface, these currents are chosen so that the
electric field outside is identical to the original problem \cite{Wriedt1998}. 
This choice of currents negates the need to compute the internal fields
which reduces the scattering problem to a system of linear equations 
relating the incident beam coefficients to the scattered beam coefficients
the relationship between each can be summarised as:
\begin{align}
	\begin{pmatrix}
		q_{mn} \\
		p_{mn} 
	\end{pmatrix}
	= \bold{T} 
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
	= \begin{pmatrix}
		T_{11} \ T_{12} \\
		T_{21} \ T_{22}
	\end{pmatrix}
	\begin{pmatrix}
		a_{mn} \\
		b_{mn}
	\end{pmatrix}
\end{align}

Where $q_{mn}$, and $p_{mn}$ are the scattering beam coefficients, and
$a_{mn}$, and $b_{mn}$ are the coefficients of the incident beam. For 
this PhD in particular we utilise an extension of the ECBM known 
as multi-sphere $T$-matrix method (\textit{mstm}), developed by Mackowski
\cite{Mackowski2011} the computational code computes the scattered field 
from each sphere within the target cluster and the incident field itself, 
this scattering between spheres converges to a final result but modern codes 
truncate the calculations to fit a desired accuracy and computational time.

The $T$-matrix method is exceptionally useful for computing the scattering 
from any arbitrary spherical aggregate. However, the $T$-matrix method by 
itself can be computationally taxing as the number of spheres increases 
\cite{Mackowski2011}. While it is possible to solve for the scattering of
each individual sphere this is only applicable for a single orientation and 
can be even slower for large aggregates \cite{Mackowski1996, Xu1995}. The 
benefit of \textit{mstm} is that the major scattering properties (scattering 
and extinction cross sections, scattering matrices, etc) can all be computed 
both for single orientations, or averaged over multiple orientations to determine
the average scattering from the target particle.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Discrete Dipole Approximation}
The discrete dipole approximation (DDA) is a general method that can 
be applied to the scattering from particles of arbitrary composition
and geometry. Developed by Purcell and Pennypacker \cite{Purcell1973}, 
the DDA method approximates the particle as being constructed of dipoles.
Each dipole interacts with both the incident field and the scattered fields
from every other dipole surrounding it. The resulting scattered field is 
identical to the scattered field produced by direct integration of 
Eq.~\ref{eq:internal} throughout the full particle volume \cite{Goedecke1988}.
The integral form for the internal electric field inside a scatterer is 
given as:
\begin{align}
	\mathbf{E(r)} = \mathbf{E_{inc}(r)} + \int_{V/V_0}d^3r\mathbf{\bar{G}(r,r')}
	\chi(\mathbf{r'})\mathbf{E(r')}
\end{align}

Where $\mathbf{\bar{G}}$ is the Greens dyadic function of free space, which
defines the impulse response between two separate dipoles; and $\chi$ is the 
susceptibility of the medium, which describes the degree of polarisation of
the medium in the presence of an electric field. With the internal field 
calculated the scattered field can be computed, one of the primary advantages
of DDA over the T-matrix method is that the the composition of the target can 
be changed freely. When comparing different computational scattering methods,
the ECBM was found to be better suited for simulating the scattering of symmetric
targets as the ECBM can use the targets symmetry to speed up calculations \cite{Wriedt1998},
however when dealing with inhomogeneous media DDA is more efficient compared to
ECBM.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Langevin Equation}
Describing any microscopic motion requires an understanding of a molecules
diffusive behaviour, for the case of optical tweezers the most complete model 
of diffusion is the Langevin equation. Models such as the Fickian, and Einstein 
derivations are sufficient for macroscopic behaviours the Langevin equation 
better describes the microscopic characteristics of any diffusive behaviour. 
The Fickian model is not suitable for the applications of optical tweezers as 
it assumes that all diffusive motion is driven by a gradient of molecular density
\cite{Gillespie2012}, not only is this often not the case but it fails to describe 
the motion of any single molecule, only the collective behaviour. Einstein's model 
however considers the collisions experienced by individual molecules \cite{Gillespie2012a}, 
if we assume that all collisions are random and only consider the behaviour 
after a finite time step $\delta t$ then the individual collisions should 
cancel out, of course trying to understand the dynamics over a shorter period
of time is not described by Einstein's model as you would have to consider
each collision in turn \cite{Gillespie2012, Gillespie2012}. This is insufficient 
for our interests as an optical tweezers relaxation time (given by 
$\tau = \kappa/\gamma$) is often so short that we must consider the individual 
collisions experienced by any one molecule. The Langevin model of diffusion 
therefore assumes that the net force on a particular particle is described 
fully by these individual collisions \cite{Gillespie2012c}:
\begin{align}
	m\frac{dv}{dt} + \gamma_0 v + F(t) = W(t)
\end{align}

Where the first term accounts for inertial forces, the second term accounts 
for friction forces which counteract the particles current motion ($\gamma_0$ 
is the friction coefficient), and the final term accounts for the random 
Brownian force. The $F(t)$ is there for convention which accounts for any 
external forces acting on the particle. We can say that the noise term 
$W(t)$ has a Gaussian probability, with a correlation function of:
\begin{align}
	&W(t) = \sqrt{2k_BT\gamma_0}\eta(t) \\
	&\langle W_i(t)W_j(t')\rangle = 2k_BT\gamma_0\delta(t-t')
\end{align}

@@

\subsection{Finite Difference}
The Finite Difference approach involves discretizing the time and spatial 
elements in order to approximate the higher order terms. If we assume that 
$x(t)$ is differentiable to n (we can find its $n^{th}$ derivative) then 
we can use the Taylor series expansion to get:
\begin{align}
	x(t+\Delta t) = x(t)+\frac{x'(t)}{1!}\Delta t + \frac{x^2(t)}{2!}\Delta t^2+...+\frac{x^n(t)}{n!}\Delta t^n+R_n(x(t))	
\end{align}
Where $R_n(x(t))$ is the remainder term between the Taylor expansion to 
term n and the actual expression. If we limit our approach to the first 
derivative only we find that for sufficiently small values of $R_1$ the 
velocity and acceleration can be approximated as:
\begin{align}
	x'(t) &\approx \frac{x(t+\Delta t)-x(t)}{\Delta t}
	\\
	x^{''}(t) &\approx \frac{x'(t+\Delta t)-x'(t)}{\Delta t} = \frac{x(t)-2x(t+\Delta t)+x(t+2\Delta t)}{\Delta t^2}
\end{align}
By reversing the time step (i.e. use $-\Delta t$) to approximate the \\
velocity and acceleration based on the previous time steps we can 
discretize the position by taking finitely small  time steps (i.e. 
$x(t) = x_i,\ x(t-\Delta t) = x_{i-1}$). The same cannot be done for 
white noise as no information is known about $W(t)$ at any time. We 
can instead say that the velocity of a Brownian particle should 
approximate our white noise as a random walk, where at each new time 
step the position changes randomly within a given range. Constricting 
the variance to $\sqrt{2D}/\Delta t$ allows us to represent the white 
noise using the finite-difference approach as:
\begin{align}
	m\frac{x_i-2x_{i-1}+x_{i-2}}{\Delta t^2} = -\gamma\frac{x_i-x_{i-1}}{\Delta t}+\sqrt{2k_BT\gamma}\frac{w_i}{\sqrt{\Delta t}}
\end{align}
Where $w_i$ is a random real number between -1 and 1, we can say that 
it is normally distributed around 0 for simulation purposes. We can 
rearrange this for $x_i$ to approximate the Brownian motion of a 
particle (setting $x_0=0$), where the characteristic time is $\tau = m/\gamma$. 
Now in the case of an optical trap the restoration time scale is 
given by $\tau_{OT}=\kappa_x/\gamma$ which for strongly trapped particles 
is far greater than the characteristic time, therefore for simulation 
purposes we can neglect the particle's inertia which allows us to 
write the motion of an optically trapped particle as:
\begin{align}
	\label{eq:sim_langevin}
	x_i = x_{i-1} + \tau_{OT}\ x_{i-1}\Delta t + \sqrt{2D\Delta t}\ w_i
\end{align}

@@

\subsection{Potential Well Analysis}
The Langevin equation for an optically trapped assumes that the trap 
acts similar to a Hookean spring that creates a potential well about 
its centre. Therefore a simple analysis method is to understand the 
height and width of said potential well.

Potential analysis is a useful technique for estimating the strength 
of an optical trap; this method assumes that the force acting on the 
particle is purely conservative, an accurate presupposition if we 
ignore the motion of the particle as it enters the trap. This is 
because the scattering force is far more significant far away from 
the potential well and is negligible if the trap strength is much 
greater than the thermal fluctuations. With this in mind we can 
write the probability of finding the particle at position $x$ as:
\begin{align}
	\frac{\rho(x)}{\rho_0} = e^{-\frac{U(x)}{k_{B}T}} 
\end{align}
which therefore means we can write the potential well as:
\begin{align}
	\label{eq:potential_well}
	U(x)=-k_BT\ ln\left(\frac{\rho(x)}{\rho_0} \right)
\end{align}
Now assuming our laser acts as a Gaussian beam we should be able 
to describe the probability distribution $\rho(x)$ centred at 
some equilibrium position $x_0$:
\begin{align}
	\label{eq:prob_dist}
	\rho(x)= \sqrt{\frac{\kappa_x}{2\pi k_BT}} exp\left(-\frac{\kappa_x}{2k_BT}(x-x_{eq})^2\right)
\end{align}

By inserting eq.~\ref{eq:prob_dist} into eq.~\ref{eq:potential_well} 
we can fit the potential well in order to determine the trap strength 
$\kappa_x$, and an estimation of the equilibrium position $x_{eq}$. 
This has some limitations in that the large fluctuations can throw 
off the fit meaning a longer acquisition time is necessary to properly 
fit the potential well, making it difficult to characterise weakly 
trapped particles who may not remain trapped for long. It also 
provides no information on the particle itself (i.e. the friction 
coefficient $\gamma$ and diffusion coefficient $D$).

@@

By taking a time average over multiple trajectories to get 
$\left<x-x_{eq}\right>$ we can get a fairly accurate estimation 
of the trap strength. Because of this requires a time average of 
the particle's displacement any large errors in the position 
measurement can have knock-on effects. Likewise with the potential 
analysis route, no information is gleaned about the particle itself.

\subsection{Mean Square Displacement}
Mean square displacement (MSD) is a common means of describing 
the random motion of a given particle (or group of particles). 
This is useful information if say for example we want to understand 
reaction kinetics on the surface of a catalyst, if we know how 
far its likely to move from the surface we can tell if its likely 
to react when a catalytic site becomes available. As it pertains 
to colloids, consider a suspension of silica spheres immersed in 
a fluid undergoing Brownian motion (as described by the Langevin 
Equation) so that:
\begin{align}
	mx''(t) + \gamma x'(t) = \eta(t)
\end{align}

Where $\gamma$ is the objects friction coefficient which for spheres 
is given as $\gamma = 6\pi\eta r$, and $\eta(t)$ is a random white 
noise variable that is directly related to the thermal energy of 
the surrounding fluid. If the motion is truly random then we should 
see an average displacement of 0 regardless of how long we measure 
for. If we wish to understand the effects of a given external factor 
(i.e. an electric field or localised heating), simply looking at 
displacement will reveal nothing of value as its difficult to 
differentiate between diffusive and a biased motion.
